Are Transformed Low Discrepancy Points Also Low Discrepancy?

### 2019-05-18 Draft_Inv.tex
Updates from Overleaf

\usepackage{psfrag,bbm,epsf}
\usepackage{amssymb, color,enumerate,float,mathtools,algorithmicx}
\usepackage{algorithm}
\usepackage[numbers]{natbib}
\usepackage{enumerate,multirow,booktabs}
\usepackage{url}
\usepackage{algpseudocode}
\usepackage{natbib}

\newcommand{\blind}{1}
\newtheorem{lem}{Lemma}
\newtheorem{thm}{Theorem}
\newtheorem{pro}{Proposition}
\newtheorem{defn}{Definition}
\newtheorem{rmk}{Remark}
\newtheorem{cor}{Corollary}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\naturals}{\mathbb{R}}
\newcommand{\vc}{\boldsymbol{c}}
\newcommand{\vx}{\boldsymbol{x}}
\newcommand{\vX}{\boldsymbol{X}}
\newcommand{\vt}{\boldsymbol{t}}
\newcommand{\vz}{\boldsymbol{z}}
\newcommand{\vu}{\boldsymbol{u}}
\newcommand{\vy}{\boldsymbol{y}}
\newcommand{\mX}{{\mathsf X}}
\newcommand{\vinfty}{\boldsymbol{\infty}}
\newcommand{\vzero}{\boldsymbol{0}}
\newcommand{\vPsi}{\boldsymbol{\Psi}}
\newcommand{\dif}{{\rm d}}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\argmax}{argmax}
\newcommand{\Udes}{\mathcal{U}}
\newcommand{\Xdes}{\mathcal{X}}
\newcommand{\FXdes}{F_{\Xdes}}
\newcommand{\cm}{\mathcal{M}}
\newcommand{\ch}{\mathcal{H}}
\newcommand{\Ftar}{F}
\newcommand{\ftar}{\varrho}
\newcommand{\cube}{\ensuremath{(0,1)^d}}
\newcommand{\bbone}{\mathbbm{1}}
\newcommand{\Ex}{\mathbb{E}}
\newcommand{\ip}[3][{}]{\ensuremath{\left \langle #2, #3 \right \rangle_{#1}}}
\newcommand{\norm}[2][{}]{\ensuremath{\left \lVert #2 \right \rVert}_{#1}}
\DeclareMathOperator{\cov}{cov}

@@

\newcommand{\FJH}[1]{{\color{blue}Fred: #1}}
\newcommand{\LK}[1]{{\color{green}Lulu: #1}} %choose your color
\newcommand{\YL}[1]{{\color{red}Yiou: #1}} %choose your color

@@

\title*{Are Transformed Low Discrepancy Points Also Low Discrepancy?}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{Yiou Li, Lulu Kang, and Fred Hickernell}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Yiou Li \at DePaul University, 1 E. Jackson
Chicago, IL 60604, \email{yli139@depaul.edu}
\and Lulu Kang \at Illinois Institute of Technology, 10 W 32nd Street, Chicago, IL 60616 \email{lkang2@iit.edu} 
\and 
Fred Hickernell \at Illinois Institute of Technology, 10 W 32nd Street, Chicago, IL 60616 \email{hickernell@iit.edu}}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

@@

Low discrepancy designs, $\Udes = \{\vu_i\}_{i=1}^N$, are typically constructed so that their empirical distribution, $F_\Udes$, approximates $F_u$, the uniform distribution on the unit cube, \cube.  
The discrepancy measures how large $F_u-F_\Udes$ is.

When the target distribution for the design, $\Ftar$, defined over the experimental domain $\Omega$, is \emph{not} the uniform distribution on the unit cube, then one typically constructs the desired design, $\Xdes$, by transforming a low discrepancy uniform design, i.e., 
\begin{equation} \label{eq:transPts}
\Xdes = \{\vx_i\}_{i=1}^N = \{\vPsi(\vu_i)\}_{i=1}^N = \vPsi(\Udes), \qquad \vPsi: \cube \to \Omega.
\end{equation}
Note that $\Ftar$ may differ from $F_u$ because $\Omega \ne \cube$ and/or $\Ftar$ is non-uniform.  
A natural transformation, $\vPsi(\vu)=[\Psi_1(u_1),\ldots,\Psi_d(u_d)]$, when $\Ftar$ has independent marginals is the inverse distribution transformation:
\begin{equation}\label{eq:inverse}
\Psi_j(u_j) = F_j^{-1}(u_j), \quad \text{where } \Ftar(\vx) = F_1(u_1) \cdots F_d(u_d).
\end{equation}
A number of transformation methods for different distributions can be found in \cite{devroye2006nonuniform}.

The question addressed in this chapter is whether the design $\Xdes$ resulting from transformation \eqref{eq:transPts} of a low discrepancy design, $\Udes$, is itself low discrepancy with respect to the target distribution $\Ftar$. 
In other words, 
\begin{equation} \label{BigQ} \tag{Q}
\text{\emph{does small $F_u - F_\Udes$ imply small $\Ftar - F_\Xdes$?}}
\end{equation}
We demonstrate that the answer may be yes or no, depending on how the question is posed.  
We discuss both cases.  
For illustrative purposes, we consider the case where $\Ftar$ is the multivariate normal distribution.

In the rest of this chapter, we first explain the discrepancy that measures how large $\Ftar - F_\Xdes$ is and motivate it from three different perspectives.  Next, we the inverse transformation can lead to inconsistency in terms the discrepancies of the designs before and after transformation. 
We also proposed a remedy method to improve the design in terms of the discrepancy between $\Ftar$ and $F_\Xdes$. 
Numerical examples are used to show the performance of the method.

@@

Our notation for discrepancy takes the form of 
\[
 D(F_{\Xdes},F,K), \ D(\Xdes,F,K), \ D(\Xdes,\varrho,K), \ D(\Xdes,\nu,K), \ D(\nu_{\Xdes},\nu,K), \text{ etc.}, 
\]
are all of which mean the same thing.  The first argument always refers to the design, the second argument always refers to the target, and the third argument is a symmetric, positive definite kernel, which is explained below.  We abuse the discrepancy notation because sometimes it is convenient to refer to the design as a set, $\Xdes$, other times by its empirical distribution, $F_{\Xdes}$, and other times by its empirical $\nu_{\Xdes}$.  Likewise sometimes it convenient to refer the target as a probability measure,  $\nu$, other times by its distribution function, $F$, and other times by its density function, $\varrho$.

In the remainder of this section we provide three interpretations of the discrepancy, summarized in Table \ref{tab:ThreeInterpTable}.  One interpretation is the norm of $\nu - \nu_{\Xdes}$.  The second and third interpretations consider the problem of evaluating the mean of a random variable $Y=f(\vX)$, or equivalently a multidimensional integral
\[
\mu = \Ex(Y) = \Ex[f(\vX)] = \int_{\Omega} f(\vx) \, \varrho(\vx) \, \dif \vx,
\]
where $\vX$ is a random vector with density $\varrho$. The second
interpretation is worst-case cubature error for integrands, $f$, in the unit ball of a Hilbert space.  The third interpretation is the root mean squared cubature error for integrands, $f$ that are stochastic processes.

\begin{table}
    \centering
     \caption{Three interpretations of the discrepancy}
     \vspace{-3ex}
\begin{equation*}
\begin{array}{c@{\qquad}ccc}
     \text{Kernel Interpretation} & \text{Discrepancy } D(\Xdes,\nu,K) = D(\Xdes,\varrho,K)\\
     \toprule \\[-1ex]
     K(\vt,\vx) = \ip[\cm]{\delta_{\vt}}{\delta_{\vx}} & \norm[\cm]{\nu - \nu_{\Xdes}} \\[2ex]
     f(\vx) = \ip[\ch]{K(\cdot,\vx)}{f} & \displaystyle \sup_{f \in \ch : \norm[\ch]{f} \le 1} \abs{\int_{\Omega} f(\vx) \, \varrho(\vx) \, \dif \vx - \frac 1N \sum_{i=1}^N f(\vx_i)}\\[4ex]
    K(\vt,\vx) = \cov\bigl(f(\vt), f(\vx) \bigr) & \displaystyle \sqrt{\Ex \abs{\int_{\Omega} f(\vx) \, \varrho(\vx) \, \dif \vx - \frac 1N \sum_{i=1}^N f(\vx_i)}^2}\\
\end{array}
\end{equation*}
    \label{tab:ThreeInterpTable}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Definition in Terms of a Norm on a Hilbert Space of Measures}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Let $(\cm, \ip[\cm]{\cdot}{\cdot})$ be a Hilbert space of measures defined on the experimental region, $\Omega$.  Assume that $\cm$ includes all Dirac measures.  Define the function $K:\Omega \times \Omega \to \reals$ in terms of inner products of Dirac measures:
\begin{equation} \label{kerDeltaDef}
    K(\vt,\vx) := \ip[\cm]{\delta_{\vt}}{\delta_{\vx}}, \qquad \forall \vt, \vx \in \Omega.
\end{equation}
It is straightforward to show that $ K$ is symmetric in its arguments and positive-definite, namely:
\begin{gather*}
K(\vx, \vt) = K(\vt, \vx) \qquad \forall \vt, \vx\in \Omega,\\
\sum\limits_{i, k=1}^N c_i c_k  K(\vx_i,\vx_k) > 0, \qquad \forall N\in\mathbb{N}, \  \vc \in\mathbb{R}^N, \ \vc \ne \vzero,  \ \Xdes \in\Omega.
\end{gather*}
For arbitrary measures $\lambda, \nu \in \cm$, their inner product can be expressed in terms of a double integral of the kernel, $K$:
\begin{equation} \label{ipMdef}
    \ip[\cm]{\lambda}{\nu} = \int_{\Omega \times \Omega} K(\vt,\vx) \, \lambda(\dif \vt) \nu(\dif \vx).
\end{equation}
This can be established directly from \eqref{kerDeltaDef} for $\cm_0$, the vector space spanned by all Dirac measures.  Letting $\cm$ be the closure of the pre-Hilbert space $\cm_0$ then yields \eqref{ipMdef}.

The discrepancy of the design $\Xdes$ with respect to the target probability measure $\nu$ using the kernel $K$, can be defined as the norm of the difference between the target probability measure, $\nu$, and the empirical probability measure for $\Xdes$:
\begin{subequations} \label{discDef}
\begin{align} 
    D^2(\Xdes,\nu,K) & := \norm[\cm]{\nu - \nu_{\Xdes}}^2 \\
    \nonumber
    & = \int_{\Omega \times \Omega} K(\vt,\vx) \, (\nu - \nu_{\Xdes})(\dif \vt) (\nu - \nu_{\Xdes})(\dif \vx) \\
    \nonumber
    & = \int_{\Omega \times \Omega} K(\vt,\vx) \, \nu(\dif \vt) \nu (\dif \vx)\\
    & \qquad \qquad - \frac 2N \sum_{i=1}^N \int_{\Omega} K(\vt,\vx_i) \, \nu(\dif \vt) + \sum_{i,k=1}^N K(\vx_i,\vx_k).
\end{align}
The formula for the discrepancy may be written equivalently in terms of the probability distribution, $F$,  or the probability density, $\varrho$, corresponding to the target probability measure, $\nu$:
\begin{align}
\nonumber
    D^2(\Xdes,F,K)  & = \int_{\Omega \times \Omega} K(\vt,\vx) \, \dif F( \vt) \dif F(\vx)\\
    & \qquad \qquad - \frac 2N \sum_{i=1}^N \int_{\Omega} K(\vt,\vx_i)  \, \dif F(\vt) + \sum_{i,k=1}^N K(\vx_i,\vx_k), \\
\nonumber
   & = \int_{\Omega \times \Omega} K(\vt,\vx) \, \varrho(\vt) \varrho (\vx) \, \dif \vt\dif \vx\\
    & \qquad \qquad - \frac 2N \sum_{i=1}^N \int_{\Omega} K(\vt,\vx_i)  \, \varrho(\vt) \, \dif \vt + \sum_{i,k=1}^N K(\vx_i,\vx_k).
\end{align}
\end{subequations}
Here, and below we abuse the notation for discrepancy, but we always preserve the meaning of the first second and third arguments.  Thus,
\[
D(\Xdes,\nu,K), \ D(\Xdes,F,K), \ D(\Xdes,\varrho,K), \ D(\nu_{\Xdes},\nu,K), \text{ etc.} 
\]
are all equivalent.

The equivalent formulas for the discrepancy in \eqref{discDef} depend inherently on the choice of $K$.  That is key to the answering our question \eqref{BigQ}.  A kernel often used is 
\begin{equation} \label{OrigKernel}
     K(\vt,\vx)  = \prod\limits_{j=1}^d\left[1+ |t_j|+ |x_j|- |x_j-t_j|\right].
\end{equation}
This kernel is plotted in Figure \ref{fig:kernelpict} for $d=1$.  Note that for this kernel in one dimension the distance between two Dirac measures is 
\begin{equation*}
    \norm[\cm]{\delta_{x} - \delta_{t}} = \sqrt{\abs{x-t}}.
\end{equation*}

\begin{figure}
    \centering
    \includegraphics[height = 4cm]{KernelPict.eps} \qquad
    \includegraphics[height = 4cm]{Kernel3DPict.eps}
    \caption{The kernel defined in \eqref{OrigKernel} for $d=1$.}
    \label{fig:kernelpict}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Definition in Terms of a Norm on a Hilbert Space of Measures}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{A Discrepancy Defined for Normal Distribution}

Consider any symmetric and positive semi-definite kernel function $K(\vx,\vt)$ on the experiment region $\Omega$. 
In other words, the kernel function satisfies 
\begin{align*}
K(\vx, \vt) &= K(\vt, \vx), \,\,\,\forall \vt, \vx\in \Omega,\\
\sum\limits_{i, k=1}^n c_ic_k & K(\vx_i,\vx_k)\geq 0, \,\,\,\forall n\in\mathbb{N^+}, c_i\in\mathbb{R}, \vx_i\in\Omega.
\end{align*}
The discrepancy of a design $P = \{\vx_i\}_{i=1}^N$ on $\Omega$ with respect to a distribution $\Ftar$ is then defined as
\begin{equation}\label{eq:Disc}
D(P) = \left\{\int_{\Omega\times\Omega} K(\vx,\vt)\dif [\Ftar(\vx)-F_{P}(\vx)]\dif [\Ftar(\vt)-F_{P}(\vt)]\right\}^{1/2},
\end{equation}
and thus, 
\begin{align}\label{eq:Disc2}\nonumber
[D(P)]^2  =& \int_{\Omega\times\Omega}  K(\vx,\vt)\dif F_{*}(\vx)\dif F_{*}(\vt) \\
- &\frac{2}{N}\sum\limits_{\vx\in P}\int_{\Omega}K(\vx,\vt)\dif F_{*}(\vt) +\frac{1}{N^2}\sum\limits_{\vx,\vx'\in P}K(\vx,\vx')
\end{align}
In \cite{Hic99a}, it is showed that, any discrepancy determined by a reproducing kernel function $K$ and a distribution $\Ftar$ is a goodness-of-fit statistic that measures how well the empirical distribution $F_P$ of design $P$ approximates the distribution $\Ftar$. 
What's more, the order of computation for evaluating the discrepancy is at worst $\mathcal{O}(N^2)$.

 Let $P = \{\vx_i\}_{i=1}^N$ be a design on $\Omega = \mathbb{R}^d$. 
 According to \eqref{eq:Disc2}, the squared discrepancy of the design $P$ with respect to the $d$-dimensional standard multivariate normal probability distribution $N({\bf 0}, {\bf I}_d)$ with cdf $\Phi$ is
\begin{eqnarray}\label{eqn:discrepancy}\nonumber
[D(P)]^2 &=& \int_{\Omega\times\Omega} K(\vx,\vt)\dif \Phi(\vx)\dif \Phi(\vt)\\
&-&\frac{2}{N}\sum_{i=1}^N \int_{\Omega}K(\vx,\vx_i)\dif \Phi(\vx)+\frac{1}{N^2}\sum_{i,k=1}^N K(\vx_i,\vx_k).
\end{eqnarray}

>>>>

\usepackage{psfrag,bbm,epsf}
\usepackage{xspace,subfigure}
\usepackage{amssymb, xcolor,enumerate,float,mathtools,algorithmicx}
\usepackage{algorithm}
\usepackage[numbers]{natbib}
\usepackage{enumerate,multirow,booktabs}
\usepackage{url}
\usepackage{algpseudocode}
\usepackage{natbib}
\usepackage{showlabels}

\newcommand{\blind}{1}
\newtheorem{lem}{Lemma}
\newtheorem{thm}{Theorem}
\newtheorem{pro}{Proposition}
\newtheorem{defn}{Definition}
\newtheorem{rmk}{Remark}
\newtheorem{cor}{Corollary}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\naturals}{\mathbb{R}}
\newcommand{\va}{\boldsymbol{a}}
\newcommand{\vb}{\boldsymbol{b}}
\newcommand{\vc}{\boldsymbol{c}}
\newcommand{\vx}{\boldsymbol{x}}
\newcommand{\vX}{\boldsymbol{X}}
\newcommand{\vt}{\boldsymbol{t}}
\newcommand{\vz}{\boldsymbol{z}}
\newcommand{\vu}{\boldsymbol{u}}
\newcommand{\vv}{\boldsymbol{v}}
\newcommand{\vy}{\boldsymbol{y}}
\newcommand{\fraku}{\mathfrak{u}}
\newcommand{\frakv}{\mathfrak{v}}
\newcommand{\mX}{{\mathsf X}}
\newcommand{\vinfty}{\boldsymbol{\infty}}
\newcommand{\vzero}{\boldsymbol{0}}
\newcommand{\vPsi}{\boldsymbol{\Psi}}
\newcommand{\dif}{{\rm d}}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\argmax}{argmax}
\newcommand{\Udes}{\mathcal{U}}
\newcommand{\Xdes}{\mathcal{X}}
\newcommand{\FXdes}{F_{\Xdes}}
\newcommand{\cm}{\mathcal{M}}
\newcommand{\ch}{\mathcal{H}}
\newcommand{\MATLAB}{\textsc{Matlab}\xspace}
\newcommand{\Ftar}{F}
\newcommand{\ftar}{\varrho}
\newcommand{\cube}{\ensuremath{(0,1)^d}}
\newcommand{\bbone}{\mathbbm{1}}
\newcommand{\Ex}{\mathbb{E}}
\newcommand{\ip}[3][{}]{\ensuremath{\left \langle #2, #3 \right \rangle_{#1}}}
\newcommand{\norm}[2][{}]{\ensuremath{\left \lVert #2 \right \rVert}_{#1}}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\err}{err}
\newcommand{\unif}{\textup{unif}}
\newcommand{\normal}{\textup{normal}}

@@

\newcommand{\FJH}[1]{{\color{blue}Fred: #1}}
\newcommand{\LK}[1]{{\color{green}Lulu: #1}} %choose your color
\newcommand{\YL}[1]{{\color{teal}Yiou: #1}} %choose your color

@@

\title*{Are Transformed Low Discrepancy Points Also Low Discrepancy?}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{Yiou Li, Lulu Kang, and Fred Hickernell}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Yiou Li \at DePaul University, 2320 N. Kenmore Avenue, Chicago, IL 60614 \email{yli139@depaul.edu}
\and Lulu Kang \at Illinois Institute of Technology, 10 W 32nd Street, Chicago, IL 60616 \email{lkang2@iit.edu} 
\and 
Fred Hickernell \at Illinois Institute of Technology, 10 W 32nd Street, Chicago, IL 60616 \email{hickernell@iit.edu}}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

@@

Low discrepancy designs, $\Udes = \{\vu_i\}_{i=1}^N$, are typically constructed so that their empirical distribution, $F_\Udes$, approximates $F_\unif$, the uniform distribution on the unit cube, \cube.  
The discrepancy measures how large $F_\unif-F_\Udes$ is.

When the target distribution for the design, $\Ftar$, defined over the experimental domain $\Omega$, is \emph{not} the uniform distribution on the unit cube, then one typically constructs the desired design, $\Xdes$, by transforming a low discrepancy uniform design, i.e., 
\begin{equation} \label{eq:transPts}
\Xdes = \{\vx_i\}_{i=1}^N = \{\vPsi(\vu_i)\}_{i=1}^N = \vPsi(\Udes), \qquad \vPsi: \cube \to \Omega.
\end{equation}
Note that $\Ftar$ may differ from $F_\unif$ because $\Omega \ne \cube$ and/or $\Ftar$ is non-uniform.  
A natural transformation, $\vPsi(\vu)=[\Psi_1(u_1),\ldots,\Psi_d(u_d)]$, when $\Ftar$ has independent marginals, is the inverse distribution transformation:
\begin{equation}\label{eq:inverse}
\Psi_j(u_j) = F_j^{-1}(u_j), \quad j =1, \ldots, d, \qquad \text{where } \Ftar(\vx) = F_1(u_1) \cdots F_d(u_d).
\end{equation}
A number of transformation methods for different distributions can be found in \cite{devroye2006nonuniform}. \FJH{How about Prof. Fang's books?}

The question addressed in this chapter is whether the design $\Xdes$ resulting from transformation \eqref{eq:transPts} of a low discrepancy design, $\Udes$, is itself low discrepancy with respect to the target distribution $\Ftar$. 
In other words, 
\begin{equation} \label{BigQ} \tag{Q}
\text{\emph{does small $F_\unif - F_\Udes$ imply small $\Ftar - F_\Xdes$?}}
\end{equation}
We demonstrate that the answer may be yes or no, depending on how the question is posed.  
We discuss both cases.  
For illustrative purposes, we consider the case where $\Ftar$ is the multivariate standard normal distribution.

In the rest of this chapter, we first explain the discrepancy that measures how large $\Ftar - F_\Xdes$ is and motivate it from three different perspectives.  Next, we show how the inverse transformation can lead to does not always preserve the low discrepancy of a design.
We also propose a remedy method to improve the design in terms of the discrepancy between $\Ftar$ and $F_\Xdes$. 
Numerical examples are used to show the performance of the method.

@@

Our notation for discrepancy takes the form of 
\[
 D(F_{\Xdes},F,K), \ D(\Xdes,F,K), \ D(\Xdes,\varrho,K), \ D(\Xdes,\nu,K), \ D(\nu_{\Xdes},\nu,K), \text{ etc.}, 
\]
all of which mean the same thing.  The first argument always refers to the design, the second argument always refers to the target, and the third argument is a symmetric, positive definite kernel, which is explained below.  We abuse the discrepancy notation because sometimes it is convenient to refer to the design as a set, $\Xdes$, other times by its empirical distribution, $F_{\Xdes}$, and other times by its empirical measure, $\nu_{\Xdes}$.  Likewise, sometimes it is convenient to refer the target as a probability measure,  $\nu$, other times by its distribution function, $F$, and other times by its density function, $\varrho$.

In the remainder of this section we provide three interpretations of the discrepancy, summarized in Table \ref{tab:ThreeInterpTable}.  These results are derived previously in various places, including \cite{Hic99a}.  One interpretation is the norm of $\nu - \nu_{\Xdes}$.  The second and third interpretations consider the problem of evaluating the mean of a random variable $Y=f(\vX)$, or equivalently a multidimensional integral
\begin{equation}\label{eq:integration}
\mu = \Ex(Y) = \Ex[f(\vX)] = \int_{\Omega} f(\vx) \, \varrho(\vx) \, \dif \vx,
\end{equation}
where $\vX$ is a random vector with density $\varrho$. The second
interpretation is worst-case cubature error for integrands, $f$, in the unit ball of a Hilbert space.  The third interpretation is the root mean squared cubature error for integrands, $f$ which are realizations of a stochastic processes.

\begin{table}
\centering
\caption{Three interpretations of the discrepancy}
\vspace{-3ex}
\begin{equation*}
\begin{array}{c@{\qquad}ccc}
\text{Kernel Interpretation} & \text{Discrepancy } D(\Xdes,\nu,K) = D(\Xdes,\varrho,K)\\
\toprule \\[-1ex]
K(\vt,\vx) = \ip[\cm]{\delta_{\vt}}{\delta_{\vx}} & \norm[\cm]{\nu - \nu_{\Xdes}} \\[2ex]
f(\vx) = \ip[\ch]{K(\cdot,\vx)}{f} & \displaystyle \sup_{f \in \ch : \norm[\ch]{f} \le 1} \abs{\int_{\Omega} f(\vx) \, \varrho(\vx) \, \dif \vx - \frac 1N \sum_{i=1}^N f(\vx_i)}\\[4ex]
K(\vt,\vx) = \cov\bigl(f(\vt), f(\vx) \bigr) & \displaystyle \sqrt{\Ex \abs{\int_{\Omega} f(\vx) \, \varrho(\vx) \, \dif \vx - \frac 1N \sum_{i=1}^N f(\vx_i)}^2}\\\bottomrule
\end{array}
\end{equation*}
\label{tab:ThreeInterpTable}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Definition in Terms of a Norm on a Hilbert Space of Measures}
\label{sec:HilbertMeasures}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Let $(\cm, \ip[\cm]{\cdot}{\cdot})$ be a Hilbert space of measures defined on the experimental region, $\Omega$.  Assume that $\cm$ includes all Dirac measures.  Define the function $K:\Omega \times \Omega \to \reals$ in terms of inner products of Dirac measures:
\begin{equation} \label{kerDeltaDef}
    K(\vt,\vx) := \ip[\cm]{\delta_{\vt}}{\delta_{\vx}}, \qquad \forall \vt, \vx \in \Omega.
\end{equation}
The squared distance between two Dirac measures in $\cm$ is then
\begin{equation} \label{distDelta}
   \norm[\cm]{\delta_{\vx} - \delta_{\vt}}^2 = K(\vt,\vt) - 2K(\vt,\vx) + K(\vx,\vx), \qquad \forall \vt, \vx \in \Omega.
\end{equation}

It is straightforward to show that $ K$ is symmetric in its arguments and positive-definite, namely:
\begin{subequations} \label{eq:sympd}
\begin{gather}
K(\vx, \vt) = K(\vt, \vx) \qquad \forall \vt, \vx\in \Omega,\\
\sum\limits_{i, k=1}^N c_i c_k  K(\vx_i,\vx_k) > 0, \qquad \forall N\in\mathbb{N}, \  \vc \in\mathbb{R}^N, \ \vc \ne \vzero,  \ \Xdes \subset \Omega.
\end{gather}
\end{subequations}
The inner product of arbitrary measures $\lambda, \nu \in \cm$ can be expressed in terms of a double integral of the kernel, $K$:
\begin{equation} \label{ipMdef}
    \ip[\cm]{\lambda}{\nu} = \int_{\Omega \times \Omega} K(\vt,\vx) \, \lambda(\dif \vt) \nu(\dif \vx).
\end{equation}
This can be established directly from \eqref{kerDeltaDef} for $\cm_0$, the vector space spanned by all Dirac measures.  Letting $\cm$ be the closure of the pre-Hilbert space $\cm_0$ then yields \eqref{ipMdef}.

The discrepancy of the design $\Xdes$ with respect to the target probability measure $\nu$ using the kernel $K$ can be defined as the norm of the difference between the target probability measure, $\nu$, and the empirical probability measure for $\Xdes$:
\begin{subequations} \label{eq:discDef}
\begin{align} 
    D^2(\Xdes,\nu,K) & := \norm[\cm]{\nu - \nu_{\Xdes}}^2 \\
    \nonumber
    & = \int_{\Omega \times \Omega} K(\vt,\vx) \, (\nu - \nu_{\Xdes})(\dif \vt) (\nu - \nu_{\Xdes})(\dif \vx) \\
    \nonumber
    & = \int_{\Omega \times \Omega} K(\vt,\vx) \, \nu(\dif \vt) \nu (\dif \vx) - \frac 2N \sum_{i=1}^N \int_{\Omega} K(\vt,\vx_i) \, \nu(\dif \vt)\\
    & \qquad \qquad  + \frac{1}{N^2} \sum_{i,k=1}^N K(\vx_i,\vx_k).
\end{align}
The formula for the discrepancy may be written equivalently in terms of the probability distribution, $F$,  or the probability density, $\varrho$, corresponding to the target probability measure, $\nu$:
\begin{align}
\nonumber
    D^2(\Xdes,F,K)  & = \int_{\Omega \times \Omega} K(\vt,\vx) \, \dif F( \vt) \dif F(\vx) - \frac 2N \sum_{i=1}^N \int_{\Omega} K(\vt,\vx_i)  \, \dif F(\vt) \\ &\qquad \qquad  + \frac{1}{N^2}  \sum_{i,k=1}^N K(\vx_i,\vx_k), \\
\nonumber
   & = \int_{\Omega \times \Omega} K(\vt,\vx) \, \varrho(\vt) \varrho (\vx) \, \dif \vt\dif \vx - \frac 2N \sum_{i=1}^N \int_{\Omega} K(\vt,\vx_i)  \, \varrho(\vt) \, \dif \vt \\
    & \qquad \qquad + \frac{1}{N^2}  \sum_{i,k=1}^N K(\vx_i,\vx_k).
\end{align}
\end{subequations}

The formulas for the discrepancy in \eqref{eq:discDef} depend inherently on the choice of $K$.  That is key to answering our question \eqref{BigQ}.  An often used kernel is 
\begin{equation} \label{eq:OrigKernel}
     K(\vt,\vx)  = \prod\limits_{j=1}^d\left[1+ \frac 12 \left(|t_j|+ |x_j|- |x_j-t_j| \right)\right].
\end{equation}
This kernel is plotted in Figure \ref{fig:kernelpict} for $d=1$.  Note that for this kernel in one dimension the distance between two Dirac measures by \eqref{distDelta} is 
\begin{equation*}
    \norm[\cm]{\delta_{x} - \delta_{t}} = \sqrt{\abs{x-t}}.
\end{equation*}

\begin{figure}
    \centering
    \includegraphics[height = 4cm]{KernelPict.eps} \qquad
    \includegraphics[height = 4cm]{Kernel3DPict.eps}
    \caption{The kernel defined in \eqref{eq:OrigKernel} for $d=1$.}
    \label{fig:kernelpict}
\end{figure}

The discrepancy for the uniform distribution on the unit cube defined in terms of the above kernel is commonly called the $L^2$ discrepancy and is expressed as
\begin{align} 
\nonumber
    D^2(\Udes,F_\unif,K)
    & = \int_{(0,1)^d \times (0,1)^d} K(\vt,\vx) \, \dif \vt \dif \vx - \frac 2N \sum_{i=1}^N \int_{(0,1)^d} K(\vt,\vu_i) \, \dif \vt\\
    \nonumber
    & \qquad \qquad  + \frac{1}{N^2} \sum_{i,k=1}^N K(\vu_i,\vu_k) \\
    \nonumber
    & = \left( \frac 43 \right)^d 
     - \frac 2N \sum_{i=1}^N \prod_{j=1}^d \left[1 + u_{ij} - \frac{u_{ij}^2}{2} \right] \\
     & \qquad \qquad + \frac 1{N^2} \sum_{i,k=1}^N \prod_{j=1}^d \left[1+ \min(u_{ij},u_{ik})\right]
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Definition in Terms of a Deterministic Cubature Error Bound}
\label{sec:DetermBound}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Now let $(\ch, \ip[\ch]{\cdot}{\cdot})$ be a reproducing kernel Hilbert space (RKHS) of functions \cite{Aro50}, $f: \Omega \rightarrow \mathbb{R}$, which appear as the integrand in \eqref{eq:integration}. By definition, the reproducing kernel, $K$, is the unique function defined on $\Omega \times \Omega$ with the properties that $K(\cdot, \vx)\in \ch$ for any $\vx \in \Omega$ and $f(\vx)=\ip[\ch]{K(\cdot,\vx)}{f}$.  This second property, implies that $K$ reproduces function values via the inner product. It can be verified that $K$ is symmetric in its arguments and positive definite as in \eqref{eq:sympd}.

The integral $\mu = \int_{\Omega} f(\vx) \, \varrho(\vx) \, \dif \vx$, which was identified as $\Ex[f(\vX)]$ in \eqref{eq:integration}, can be approximated by a sample mean, 
\begin{equation}\label{eq:cubature}
\hat{\mu}=\frac{1}{N}\sum_{i=1}^N f(\vx_i).
\end{equation}
The quality of this approximation to the integral, or cubature, depends in part on how well the empirical distribution of the design, $\Xdes = \{\vx_i\}_{i=1}^N$, matches the target distribution $F$ associated with the density function $\varrho$.

Define the cubature error as  
\begin{align}
\nonumber 
\err(f,\Xdes) & = \mu - \hat{\mu} = \int_{\Omega} f(\vx)\, \varrho(\vx) \dif \vx -\frac{1}{N}\sum_{i=1}^N f(\vx_i) \\
& =\int_\Omega f(\vx) \, \dif[F(\vx)-F_{\Xdes}(\vx)]. \label{eq:err}
\end{align}
Under modest assumptions on the reproducing kernel,  $\err(\cdot, \Xdes)$ is a bounded, linear functional, so. 
by the Riesz representation theorem, there exists a unique representer, $\xi \in \ch$, such that 
\[
\err(f, \Xdes)=\ip[\ch]{f}{\xi}, \quad \forall f\in \ch.
\]
The reproducing kernel allows us to write down an explicit formula for that representer, namely, $\xi(\vx)=\ip[\ch]{K(\cdot,\vx)}{\xi}=\err(K(\cdot,\vx),\Xdes)$.
By the Cauchy-Schwarz inequality, there is a tight bound on the squared cubature error, namely
\begin{equation}\label{eq:wcErrA}
\abs{\err(f,\Xdes)}^2=\ip[\ch]{f}{\xi}^2\leq \norm[\ch]{f}^2 \norm[\ch]{\xi}^2.
\end{equation}
The first term on the right describes the contribution to the cubature error made by the nature of the integrand, while the second term describes the contribution made by the quality of the cubature rule.
